pSEOBNRv4HM

### 2020-11-09 papers/intro_paper/intro_paper.tex
more on waveform model

In the observer's frame, the GW polarizations can be written as 
$h_+(\iota,\varphi;t ) - i h_\times(\iota,\varphi;t) = \sum_{\ell, m} {}_{-\!2}Y_{\ell m}(\iota,\varphi)\, h_{\ell m}(t)$, where $\iota$ denotes the inclination angle computed with respect to the direction perpendicular to the orbital plane, $\varphi$ is the azimuthal direction to the observer, ${}_{-\!2}Y_{\ell m}(\theta,\phi)$ are the $-2$ spin-weighted spherical harmonics. The SEOBNR model we employ includes the $(\ell, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$ modes. For each $(\ell, m)$, the inspiral-(plunge-)merger-ringdown SEOBNR waveform is schematically given by $h_{\ell m}(t) = h_{\ell m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{\ell m} - t) + h_{\ell m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{
    match}^{\ell m})$, where $\theta(t)$ is the Heaviside step function, $h_{\ell m}^\mathrm{insp-plunge}$ represents the inspiral-plunge part of the waveform, whereas $h_{\ell m}^\mathrm{merger-RD}$ denotes the merger-ringdown phase. 
  %  
The merger-ringdown SEOBNR modes read~\citep{Bohe:2016gbl,Cotesta:2018fcv}
%
\begin{equation}
  \label{RD}
  h_{\ell m}^\mathrm{merger-RD}(t) = \nu A_{\ell m}(t) e^{i\phi_{\ell m} (t)}\,
  e^{-i\sigma_{\ell m0} (t-t_\mathrm{match}^{\ell m})} \quad t \geq t_{\rm match}^{\ell m}\,,
\end{equation}
%
where $\nu$ is the symmetric mass ratio of the binary and $\sigma_{\ell m0} = \omega_{\ell m 0} -i/\tau_{\ell m 0}$ denotes the complex frequency of the fundamental QNMs of the remnant BH. We denote the oscillation frequencies by $\omega_{\ell m  0}\equiv \Re(\sigma_{\ell m0})$ and the decay times by $\tau_{\ell m 0}\equiv -1/\Im(\sigma_{\ell m0})$. The explicit expressions for the amplitudes $A_{\ell m}(t)$ and phases $\phi_{\ell m} (t)$ can be found in Ref.~\citep{Cotesta:2018fcv}, and are chosen such that the functions $A_{\ell  m}$ and $\phi_{\ell  m}$ are of class $C^1$ at $t=  t_{\rm match}^{\ell m}$, when matching the merger-ringdown waveform to the inspiral-plunge SEOBNR waveform $h_{\ell m}^\mathrm{inspiral-plunge}(t)$.

In the SEOBNR model constructed in Ref.~\cite{Cotesta:2018fcv}, the complex frequencies $\sigma_{\ell m 0}$  are expressed in terms of the final BH mass and spin~\cite{Berti:2005ys,Berti:2009kk}, and the latter are related to the BBH's component masses and spins through NR--fitting-formulas obtained in GR~\cite{Taracchini:2013rva,Hofmann:2016yih}. Here instead, in the spirit of what was done in \paperone, we promote the QNM (complex) frequencies to be free parameters of the model, while keeping the inspiral-plunge modes $h_{\ell m}^\mathrm{inspiral-plunge}(t)$ fixed to their GR values (henceforth, pSEOBNR model). We note that when leaving $\sigma_{\ell m}$ to vary freely, the amplitudes $A_{\ell m}(t)$ and phases $\phi_{\ell m} (t)$ will in general also differ from the GR prediction, since those functions depend on the QNM complex frequencies (see~\citep{Cotesta:2018fcv}).

For the specific applications of this paper, we will normally only allow either $\sigma_{220}$ or $\sigma_{330}$ to vary, while keeping all the other modes to coincide with GR. However the model can be easily used to measure any QNM complex frequency available in the waveform.

>>>>

In the observer's frame, the GW polarizations can be written as 
%
\begin{equation}
h_+(\iota,\varphi;t ) - i h_\times(\iota,\varphi;t) = \sum_{\ell, m} {}_{-\!2}Y_{\ell m}(\iota,\varphi)\, h_{\ell m}(t)\,,
\end{equation}
%
where $\iota$ denotes the inclination angle computed with respect to the direction perpendicular to the orbital plane, $\varphi$ is the azimuthal direction to the observer, ${}_{-\!2}Y_{\ell m}(\theta,\phi)$ are the $-2$ spin-weighted spherical harmonics. The SEOBNR model we employ includes the $(\ell, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$ modes~\cite{Cotesta:2018fcv}. For each $(\ell, m)$, the inspiral-(plunge-)merger-ringdown SEOBNR waveform is schematically given by
%
\begin{equation}
h_{\ell m}(t) = h_{\ell m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{\ell m} - t) + h_{\ell m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{match}^{\ell m})\,,
\end{equation}
where $\theta(t)$ is the Heaviside step function, $h_{\ell m}^\mathrm{insp-plunge}$ represents the inspiral-plunge part of the waveform, whereas $h_{\ell m}^\mathrm{merger-RD}$ denotes the merger-ringdown phase.   
The merger-ringdown SEOBNR modes read~\citep{Bohe:2016gbl,Cotesta:2018fcv}
%
\begin{equation}
\label{RD}
h_{\ell m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{\ell m}(t)\ e^{i \tilde{\phi}_{\ell m}(t)} \ e^{-i \sigma_{\ell m 0}(t-t_{\textrm{match}}^{\ell m})},
\end{equation}
%
where $\nu$ is the symmetric mass ratio of the binary and $\sigma_{\ell m0} = 2\pi f_{\ell m 0} -i/\tau_{\ell m 0}$ denotes the complex frequency of the fundamental QNMs of the remnant BH. We denote the oscillation frequencies by $f_{\ell m  0}\equiv \Re(\sigma_{\ell m0})/(2\pi)$ and the decay times by $\tau_{\ell m 0}\equiv -1/\Im(\sigma_{\ell m0}) $. 
The functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ are given by~\cite{Bohe:2016gbl,Cotesta:2018fcv}:
%
\begin{equation}
\label{eq:ansatz_amp}
\tilde{A}_{\ell m}(t) = c_{1,c}^{\ell m} \tanh[c_{1,f}^{\ell m}\ (t-t_{\textrm{match}}^{\ell m}) \ +\ c_{2,f}^{\ell m}] \ + \ c_{2,c}^{\ell m},
\end{equation}
%
\begin{equation}
\label{eq:ansatz_phase}
\tilde{\phi}_{\ell m}(t) = \phi_{\textrm{match}}^{\ell m} - d_{1,c}^{\ell m} \log\left[\frac{1+d_{2,f}^{\ell m} e^{-d_{1,f}^{\ell m}(t-t_{\textrm{match}}^{\ell m})}}{1+d_{2,f}^{\ell m}}\right],
\end{equation}
%
where $ \phi_{\textrm{match}}^{\ell m}$ is the phase of the inspiral-plunge mode $(\ell, m)$ computed at $t = t_{\textrm{match}}^{\ell m}$. The coefficients $d_{1,c}^{\ell m}$ and $c_{i,c}^{\ell m}$ with $i = 1,2$
are fixed by imposing that the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ are of class $C^1$ at $t = t_{\textrm{match}}^{\ell m}$, when matching the merger-ringdown waveform to the inspiral-plunge SEOBNR waveform $h_{\ell m}^\mathrm{inspiral-plunge}(t)$. This allows to write the coefficients $c_{i,c}^{\ell m}$ as~\cite{Cotesta:2018fcv}:
% in terms of $c_{1,f}^{\ell   m},\ c_{2,f}^{\ell m},\ \sigma^\textrm{R}_{\ell m},\ |h_{\ell    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell  m})|,\ \partial_t|h_{\ell    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell m})|$ as
\begin{align} 
c_{1,c}^{\ell m} &= \frac{1}{c_{1,f}^{\ell
    m} \nu} \big[ \partial_t|h_{\ell
    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell m})| \nonumber \\
    &- \sigma^\textrm{R}_{\ell m} |h_{\ell
    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell
    m})|\big] \cosh^2{(c_{2,f}^{\ell m})}, \\
c_{2,c}^{\ell m} &= -\frac{ |h_{\ell
    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell
    m})|}{\nu} + \frac{1}{c_{1,f}^{\ell
    m} \nu} \big[ \partial_t|h_{\ell
    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell m})|  \nonumber \\
    &- \sigma^\textrm{R}_{\ell m} |h_{\ell
    m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell
    m})|\big] \cosh{(c_{2,f}^{\ell m})}\sinh{(c_{2,f}^{\ell m})}, \\ \nonumber   
\end{align}
and $d_{1,c}^{\ell m}$ as
\begin{align}    
d_{1,c}^{\ell m} &= \left[\omega_{\ell m}^{\textrm{insp-plunge}}(t_{\textrm{match}}^{\ell m}) -  \sigma^\textrm{I}_{\ell
      m}\right]\frac{1+ d_{2,f}^{\ell m}}{d_{1,f}^{\ell m}d_{2,f}^{\ell m}}\,,
\end{align}
%
where we denoted $\sigma_{\ell m}^\textrm{R} \equiv \Im (\sigma_{\ell m0}) < 0$ and  $\sigma_{\ell m}^\textrm{I} \equiv -\Re (\sigma_{\ell m0})$, and $\omega_{\ell m}^{\textrm{insp-plunge}}(t)$ is the frequency of the inspiral-plunge EOB mode. The coefficients $c_{i,f}^{\ell m}$ and $d_{i,f}^{\ell m}$ are obtained through fits to NR and 
Teukolsky-equation--based waveforms and can be found in Appendix C of Ref.~\cite{Cotesta:2018fcv}.

In the SEOBNR model constructed in Ref.~\cite{Cotesta:2018fcv}, the complex frequencies $\sigma_{\ell m 0}$  are expressed in terms of the final BH mass and spin~\cite{Berti:2005ys,Berti:2009kk}, and the latter are related to the BBH's component masses and spins through NR--fitting-formulas obtained in GR~\cite{Taracchini:2013rva,Hofmann:2016yih}. Here instead, in the spirit of what was done in \paperone, we promote the QNM (complex) frequencies to be free parameters of the model, while keeping the inspiral-plunge modes $h_{\ell m}^\mathrm{inspiral-plunge}(t)$ fixed to their GR values (henceforth, pSEOBNR model). 
We note that when leaving $\sigma_{\ell m}$ to vary freely, the functions $\tilde{A}_{\ell m}(t)$ and $\tilde{\phi}_{\ell m}(t)$ will in general also differ from the GR prediction, since those functions depend on the QNM complex frequencies, as can be seen from the expressions for $c_{i,c}^{\ell m}$ and $d_{1,c}^{\ell m}$.

For the specific applications of this paper, we will normally only allow either $\sigma_{220}$ and/or $\sigma_{330}$ to vary, while keeping all the other modes to coincide with GR. However the model can be easily used to measure any QNM complex frequency available in the waveform.
